Fast, Scalable Phrase-Based SMT Decoding

### 2016-05-05 acl.2016/acl2016.tex
changes after review

~\newcite{mfernandez2016boosting} was concerned with multi-core speed but treat the decoding process as a black box within a parallelization framework.

@@

The lexicalized reordering model assign a probability to a translation rule, given the relative ordering of the rule in the hypothesis.

We integrate the lexicalized model file into the translation model but storing the model's probabilities in the phrase-table. This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{lex-ro.png}} 
\end{tabular}
\caption{Decoding time with Compact Lexicalized Reordering, and integrated into a model the phrase-table}
\label{fig:lex-ro}
\end{figure} 
Integrating the lexicalized reordering model into the translation model decreases decoding time by 29\% with a single core but it is over 5 times faster using all cores. In fact, the decoding time with the integrated model is similar to that \emph{without} a lexilized reordering model. Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores. This is in contrast to using a separate lexicalized reordering model where decoding time flatten out and actually worsens after approximately 15 threads.

@@

Figure~\ref{fig:speed} shows decoding speed against the number of threads, measured in words translated per second. There is a constant increase in decoding speed when more threads are used, only decreasing slightly after 16 threads when hyperthreads are used. Overall, decoding is 12.5 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.

@@

Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores. When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding. It is also 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.

@@

>>>>

~\newcite{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.

@@

The lexicalized reordering model requires a probability distribution of the reordering behaviour of each translation rule learnt from the training data. This is represented in the model file as a fixed number of probabilities for each rule, exactly how many probabilities is dependant on the model's parameterization during training. During decoding, a probability from this distribution is assigned to each hypothesis according to the reordering of the translation rule.

Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized reordering model distributions for each rule. During decoding, the probability distribution is then taken from the translation model instead of querying a separate file.

This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{lex-ro.png}} 
\end{tabular}
\caption{Decoding time with Compact Lexicalized Reordering, and integrated into a model the phrase-table}
\label{fig:lex-ro}
\end{figure} 
Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads. In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized reordering model\footnote{Strangely, using lexicalized reordering model results in slightly faster decoding than without the model. We double checked with other datasets. We suspect this may be an artefact of the algorithm or data structures, for example, more work is needed to resolve hypothesis recombination.}. Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores.

@@

Showing decoding \emph{speed} rather than decoding time better illustrate the scalability of our approach. Figure~\ref{fig:speed} shows decoding speed against the number of threads, measured in words translated per second. There is a constant increase in decoding speed when more threads are used, decreasing slightly after 16 threads when virtual cores are employed by the CPU. Overall, decoding is 12.5 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.

@@

@@

%Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores. When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding.

Our work is 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.
